I.Sicily. An EpiDoc corpus for ancient Sicily

### 2016-01-21 source/31/EAGLE2016_I Sicily_final copy.tex
Paper 31 (Prag) edited and copied in the pdfs folder as EAGLE_2016_submission_31.pdf

\noindent Sicily is traditionally described as a `melting pot', the `crossroads of the Mediterranean'. The negative consequences
of the separation of epigraphic material according to linguistic traditions have recently been highlighted and directly
confronted by the \emph{Corpus Inscriptionum Iudaeae/Palaestinae} (\emph{CIIP}), edited by H. Cotton et al.\footnote{ Original notice
in \citet{cotton_corpus_1999}; presentation in \citet{cotton_corpus_2010}.} \emph{I.Sicily} sets out to follow in that mould, since the
different linguistic traditions of Sicily not only exist side-by-side but interact constantly throughout the island’s
history, and no study of the epigraphic material can afford to ignore contemporary and parallel material in the other
languages.\footnote{ See e.g. \citet{manganaro_greco_1993}, \citet{prag_epigraphy_2002}, \citet{salmeri_i_2004}, \citet{korhonen_language_2011}, \citet{tribulato_language_2012}.} The situation
created by basic technologies such as Unicode and EpiDoc XML mean that there is now no reason not to be language
agnostic in the inclusion of material (the point may be obvious, but the tendency towards language-specific \emph{corpora} is
still marked). The opportunities and possibilities offered by these technologies are considerable, even at the most
basic level, since, for example, searching can be made language specific or language neutral. One obvious area where
Sicilian studies are currently hampered by this partitioning is in the study of onomastics. The \emph{Lexicon of Greek
Personal Names} records most instances of Greek names for the island, but Sicily is no less rich in non-Greek names
(Latin and others), and at present there is no \emph{onomasticon} for the island.\footnote{ \url{www.lgpn.ox.ac.uk/} [accessed
26.09.2015].} Simply by the marking-up and indexing of all names in the island’s inscriptions, \emph{I.Sicily} will generate a
powerful tool for future study. Although \emph{I.Sicily} in its first phase is not undertaking morphological or syntactical
mark-up, the encoding of all these text in XML constitutes the necessary first stage in such a development, and we see
this as a highly desirable future project, and the possibilities for the field of historical linguistics are
considerable. The incorporation of a full range of metadata on the epigraphic support, geographical location,
chronology, etc. will likewise allow detailed analysis of cultural patterns and their relationship to language-use over
time.\footnote{ See \citet{prag_epigraphy_2002} for a first effort in this direction.}

@@

\noindent Sicily presents a particularly extreme version of the common problem of identifying a text and its publication record.
No existing \emph{corpus} in either Greek or Latin comes close to full coverage (\emph{CIL} X and \emph{IG} XIV are the largest individual
traditional \emph{corpora} for the region, but both are over 125 years old and cover less than 30\% of the material now
known).\footnote{\citet{_inscriptiones_1883}; \citet{kaibel_inscriptiones_1890}.} Existing online databases improve on this situation, but the results
obtainable are of very varied value. The most comprehensive, in terms of the range of data recorded, is \emph{EDR} (with which
\emph{I.Sicily} is collaborating), which currently reports 1,906 records for `Sicilia'; but this reduces to 833 when limited
to texts on stone (`lapis' or `marmor'); contrast \emph{I.Sicily}, with 2,563 records at the time of writing.\footnote{
\url{www.edr-edr.it/} [accessed 26.09.2015]} \emph{Clauss Slaby} reports 4,374 records for `Sicilia' (including Christian
inscriptions, excluding `sigilla impressa'), but the return is inclusive of all kinds of epigraphic material, without
indication or discrimination, contains some duplication, is much harder to reconcile to existing records, and records
only text.\footnote{ \url{www.manfredclauss.de/} [accessed 26.09.2015].} The \emph{PHI} database of Greek inscriptions has a rich
record of published Greek texts, but is text only and limited in outputs.\footnote{
\url{http://noapplet.epigraphy.packhum.org/regions/1156} [accessed 26.09.2015].} \emph{SEG} references are available for 733
inscriptions on stone and \emph{AE} references for 328 (data taken from the \emph{I.Sicily} database and based upon comprehensive
manual trawls of \emph{SEG} and \emph{AE}).

>>>>

\noindent Sicily is traditionally described as a `melting pot', the `crossroads of the Mediterranean'. The negative consequences
of the separation of epigraphic material according to linguistic traditions have recently been highlighted and directly
confronted by the \emph{Corpus Inscriptionum Iudaeae/Palaestinae} (\emph{CIIP}), edited by H. Cotton et al.\footnote{ Original notice
in \citet{cotton_corpus_1999}; presentation in \citet{Cotton:2007aa}.} \emph{I.Sicily} sets out to follow in that mould, since the
different linguistic traditions of Sicily not only exist side-by-side but interact constantly throughout the island’s
history, and no study of the epigraphic material can afford to ignore contemporary and parallel material in the other
languages.\footnote{ See e.g. \citet{manganaro_greco_1993}, \citet{prag_epigraphy_2002}, \citet{salmeri_i_2004}, \citet{korhonen_language_2011}, \citet{tribulato_language_2012}.} The situation
created by basic technologies such as Unicode and EpiDoc XML mean that there is now no reason not to be language
agnostic in the inclusion of material (the point may be obvious, but the tendency towards language-specific \emph{corpora} is
still marked). The opportunities and possibilities offered by these technologies are considerable, even at the most
basic level, since, for example, searching can be made language specific or language neutral. One obvious area where
Sicilian studies are currently hampered by this partitioning is in the study of onomastics. The \emph{Lexicon of Greek
Personal Names} records most instances of Greek names for the island, but Sicily is no less rich in non-Greek names
(Latin and others), and at present there is no \emph{onomasticon} for the island.\footnote{ \url{www.lgpn.ox.ac.uk/} [accessed
26.09.2015].} Simply by the marking-up and indexing of all names in the island’s inscriptions, \emph{I.Sicily} will generate a
powerful tool for future study. Although \emph{I.Sicily} in its first phase is not undertaking morphological or syntactical
mark-up, the encoding of all these text in XML constitutes the necessary first stage in such a development, and we see
this as a highly desirable future project, and the possibilities for the field of historical linguistics are
considerable. The incorporation of a full range of metadata on the epigraphic support, geographical location,
chronology, etc. will likewise allow detailed analysis of cultural patterns and their relationship to language-use over
time.\footnote{ See \citet{prag_epigraphy_2002} for a first effort in this direction.}

@@

\noindent Sicily presents a particularly extreme version of the common problem of identifying a text and its publication record.
No existing \emph{corpus} in either Greek or Latin comes close to full coverage (\emph{CIL} X and \emph{IG} XIV are the largest individual
traditional \emph{corpora} for the region, but both are over 125 years old and cover less than 30\% of the material now
known).\footnote{\citet{Mommsen:1883aa}; \citet{kaibel_inscriptiones_1890}.} Existing online databases improve on this situation, but the results
obtainable are of very varied value. The most comprehensive, in terms of the range of data recorded, is \emph{EDR} (with which
\emph{I.Sicily} is collaborating), which currently reports 1,906 records for `Sicilia'; but this reduces to 833 when limited
to texts on stone (`lapis' or `marmor'); contrast \emph{I.Sicily}, with 2,563 records at the time of writing.\footnote{
\url{www.edr-edr.it/} [accessed 26.09.2015]} \emph{Clauss Slaby} reports 4,374 records for `Sicilia' (including Christian
inscriptions, excluding `sigilla impressa'), but the return is inclusive of all kinds of epigraphic material, without
indication or discrimination, contains some duplication, is much harder to reconcile to existing records, and records
only text.\footnote{ \url{www.manfredclauss.de/} [accessed 26.09.2015].} The \emph{PHI} database of Greek inscriptions has a rich
record of published Greek texts, but is text only and limited in outputs.\footnote{
\url{http://noapplet.epigraphy.packhum.org/regions/1156} [accessed 26.09.2015].} \emph{SEG} references are available for 733
inscriptions on stone and \emph{AE} references for 328 (data taken from the \emph{I.Sicily} database and based upon comprehensive
manual trawls of \emph{SEG} and \emph{AE}).
